pSEOBNRv4HM

### 2020-09-08 papers/intro_paper/intro_paper.tex
updated tex

@@

\abhi{This section will basically provide an overview of Bayesian inference. Starting from the generalised definition of likelihood, provide an expression for the posterior on the delta QNM parameters.}

@@

\abhi{A small sub-section dedicated to priors used in this analysis. Very common question from referees.}

@@

\abhi{The GW150914-like injections we have:}
\begin{itemize}
\item{an SEOBNRv4HM software injection with the 22 mode frequencies estimated}
\item{an SEOBNRv4HM software injection with the 33 mode frequencies estimated}
\item{a non-spinning NR injection with parameters close to the real event but higher SNR for which 22 mode frequencies are estimated}
\item{a non-spinning highly-asymmetric (q=6) NR injection with total mass similar to GW150914 and high SNR for which 22 and 33 mode frequencies are estimated}
\end{itemize}

@@

\abhi{Results obtained as part of the review on GW150914 and other O1-O2 events which are of relevance}

@@

>>>>

\newcommand{\h}{\mathpzc{h}}
\newcommand{\Hhat}{\hat{\mathpzc{H}}}
\newcommand{\B}{\mathpzc{B}}
\newcommand{\hlm}{\mathpzc{h}_{\ell m}}
\newcommand{\xilm}{\xi_{\ell m}}
\newcommand{\Ylm}{{Y}^{-2}_{\ell m}}
\newcommand{\Y}{{Y}^{-2}}
\newcommand{\hc}{h_\times}
\newcommand{\hp}{h_+}
\newcommand{\Fc}{F_\times}
\newcommand{\Fp}{F_+}
\newcommand{\Mf}{M_f}
\newcommand{\cA}{\mathpzc{A}}
\newcommand{\lm}{_{\ell m}}
\newcommand{\deff}{d_\mathrm{eff}}
\newcommand{\rmi}{\mathrm{i}}
\newcommand{\blambda}{\bm{\lambda}}
\newcommand{\btheta}{\bm{\theta}}
\newcommand{\bxi}{\bm{\xi}}
\newcommand{\bzeta}{\bm{\zeta}}
\newcommand{\bs}[1]{\bm{\vec{s}_{#1}}}
\newcommand{\Mo}{M_{\odot}}
\newcommand{\FFe}{\mathrm{FF}_\mathrm{eff}}
\newcommand{\FF}{\mathrm{FF}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\rhoopt}{\rho_\mathrm{opt}}
\newcommand{\rhosubopt}{\rho_\mathrm{subopt}}
\newcommand{\fqnm}{f}
\newcommand{\sigmaqnm}{\sigma}
\newcommand{\n}{\mathbf{n}}
\newcommand*{\skymapscale}{0.5}
\newcommand*{\paramestscale}{0.455}

@@

A gravitational waveform from the (quasi-circular) coalescence of two black holes is completely described in GR by a set \textit{intrinsic} parameters, $\bxi$: the masses, $m_1, m_2$ and spins, $\bs1, \bs2$ of the initial binary and a reference time and phase $t_c, \phi_c$ respectively; and a set of \textit{extrinsic} parameters, $\btheta$: the location of the binary given by the right ascension, $\alpha$, declination,$\delta$ and the luminosity distance, $d_L$, and its orientation described through the inclination of the binary $\iota$ and its polarisation $\psi$. Also allowing for additional degrees of freedom $\bzeta$ like the complex QNM frequencies described in section \ref{sec:model}, one can then use a Bayesian formalism to obtain a posterior probability density function on the combined parameter set $\blambda = \{\bxi, \btheta, \bzeta\}$:

\begin{equation}
p(\blambda | d, H, I) \propto P(\blambda | H,I) \mathcal{L}(d | \blambda, H, I)
\end{equation}

where $P(\blambda | H,I)$ is the prior probability density function and $\mathcal{L}(d | \blambda, H, I)$ is the likelihood of obtaining the data for a given choice of the parameter set $\blambda$. We assume the hypothesis $H$ to be our waveform model described in section \ref{sec:model} and $I$ to be any other information assumed during the analysis.

Given the full-dimensional posterior probability density function $p(\blambda | d, H, I)$, we can marginalise over the \emph{nuisance} parameters, to obtain the marginalised posterior probability density function on our QNM parameters $\bzeta$, as:

\begin{equation}
p(\bzeta | d, H, I) = \int p(\blambda | d, H, I) d\bxi d\btheta
\end{equation}

Throughout our analysis, we assume the noise in the detector to be well-approximated by the Gaussian random process with a fixed power spectral density. This allows us to define a Gaussian likelihood function $\mathcal{L}(d | \blambda, H, I)$, which we stochastically sample over using either a Markhov-Chain-Monte-Carlo based algorithm implemented in the LIGO Algorithms Library.

@@

Throughout our analysis we assume a completely uninformed prior in the component masses $m_1, m_2$, uniformly distributed in the range (XX,YY). Our prior on the spins are uniform in the magnitude between (0,1) and isotropic in spin orientation. Our prior on the distance varies as $d_L^2$ upto a distance of XX. For the rest of the parameters we use standard priors as defined in the documentation (CITE Vietch et al. 2015 LALInference paper). For our non-GR ringdown parameters, we assume priors uniform in the range (XX,YY). For $d\tau = -1$, we encounter a singularity (the imaginary part of the frequency goes to infinity), which we avoid by restricting the minimum of the prior on $d\tau$ to be greater than $-1$.

@@

\begin{figure*}
	\includegraphics[width=\textwidth]{figures/GW150914_swinj_GR_vs_modGR.png}
	\caption{final figure will contain a third set of contours for modGR recovered assuming GR to show bias. That run is ongoing.}
\end{figure*}

\abhi{The GW150914-like injections we have:}
\begin{itemize}
\item{an SEOBNRv4HM software injection with the 22 mode frequencies estimated}
\item{a non-spinning NR injection with parameters close to the real event but higher SNR for which 22 mode frequencies are estimated}
\item{a non-spinning highly-asymmetric (q=6) NR injection with total mass similar to GW150914 and high SNR for which 22 and 33 mode frequencies are estimated}
\end{itemize}

@@

\begin{figure}
	\includegraphics[width=0.4\textwidth]{figures/GW150914.png}
	\caption{include contours from pyRING}
\end{figure}

\abhi{Results obtained as part of the review on GW150914 and other O1-O2 events which are of relevance}
\abhi{A plot showing the fractional quantities}
\abhi{A table with absolute quantites similar to GWTC-2 TGR}

\begin{tabular}{llllllll}
\toprule
Event & \multicolumn{3}{c}{Redshifted} & \hphantom{X} & \multicolumn{3}{c}{Redshifted} \\
& \multicolumn{3}{c}{frequency [Hz]} & \hphantom{X} & \multicolumn{3}{c}{damping time [ms]} \\[0.075cm]
\cline{2-4}
\cline{6-8}
& IMR & DS & pSEOB & \hphantom{X} & IMR & DS & pSEOB \\
\midrule

GW150914 &
$249^{+9}_{-7}$ &
$247^{+14}_{-16}$ &
$-$ &
\hphantom{X} &
$4.1^{+0.3}_{-0.2}$ &
$4.8^{+3.7}_{-1.9}$ &
$-$
\\[0.075cm]

GW170104 &
$286^{+16}_{-27}$ &
$228^{+71}_{-102}$ &
$-$ &
\hphantom{X} &
$3.5^{+0.4}_{-0.3}$ &
$3.6^{+36.2}_{-2.1}$ &
$-$
\\[0.075cm]

GW170729 &
$161^{+13}_{-14}$ &
$148^{+11}_{-11}$ &
$-$ &
\hphantom{X} &
$7.8^{+1.8}_{-1.5}$ &
$10.4^{+8.2}_{-4.7}$ &
$-$
\\[0.075cm]

GW170814 &
$295^{+13}_{-12}$ &
$527^{+340}_{-332}$ &
$-$ &
\hphantom{X} &
$3.7^{+0.3}_{-0.2}$ &
$25.1^{+22.2}_{-19.0}$ &
$-$
\\[0.075cm]

GW170823 &
$199^{+14}_{-16}$ &
$222^{+664}_{-62}$ &
$-$ &
\hphantom{X} &
$5.4^{+1.0}_{-0.6}$ &
$13.4^{+31.8}_{-9.8}$ &
$-$
\\[0.075cm]

\bottomrule
\end{tabular}

@@

\begin{figure}[H]
	\includegraphics[width=0.4\textwidth]{figures/S190521g_noiseinjections.png}
	\caption{a single figure for both Gaussian and Real noise injections, color coded, to save space}
\end{figure}

We injected NRSur7dq4 signal using the MaxL parameters from EXP30, in 2.5 hours of data surrounding the event at: t0-1.5 hours, t0-1., t0-0.5, t0+0.5, t0+1. For 3 of the 5 noise realisations, corresponding to t0-1 hour, t0+0.5 hours, and t0+1.0 hour we recover a damping time similar to the actual event, whereas for a third case, t0-0.5 hours, we obtain a damping time consistent at the 90\% credible interval. L1-SNR goes down by more than 3 in some runs (maximum L1-SNR 12, minimum 9), indicative of how the noise variability is strongly affecting the recovered parameters. For some injections the mass ratio is biased, although final spin (most sensitive to mass ratio) is not affected much.  domega varies quite a lot among runs, instead dtau seems well behaved around zero. Effective frequency is a stable parameter (always well measured), while effective tau is much more sensitive to the noise configuration.
